Physical motivation for differentiability

### 2021-08-24 Overviews/Differentiability/Differentiability.tex
Differentiability: clarifying continuity of functional

@@

@@

\subsubsection{Smoothness as convenient to express approximations of linear functionals}

\begin{defn}
	Let $X$ be a manifold of dimension $n$. A \textbf{k-surface} is a subset $U \subseteq X$ that is a $k$-dimensional manifold (with boundary) with the subspace topology. We note $S^k(X)$, or simply $S^k$, the set of all $k$-surfaces and $S(X) = \bigcup_{i=0}^{n} S^k$ the set of all possible $k$-surfaces of all dimensions.
\end{defn}

\begin{defn}
	A \textbf{k-functional} is a function $f_k : S^k \to \mathbb{R}$ with the following properties:
	\begin{description}
		\item[Linear] $f_k(\sigma_1 \cup \sigma_2) = f_k(\sigma_1) + f_k(\sigma_2)$ for every $\sigma_1, \sigma_2 \in S^k$ such that $\sigma_1 \cap \sigma_2 = \emptyset$
		\item[No contribution from boundary] $f_k(\sigma) = f_k(\sigma \setminus \partial \sigma)$ for every $\sigma \in S^k$
		\item[Commutes with the limit] $\lim\limits_{i \to \infty} f_k(\sigma_i^k) = f_k(\lim\limits_{i \to \infty}\sigma_i^k)$
	\end{description}
\end{defn}

\begin{desid}
	Let $X$ be a differentiable manifold. For every $k$-functional $f_k$ there exists a suitable $k$-form $\omega_k$ whose integral $\int_U \omega_k$ in a region $U$ approximates $f_k(U)$ at a desired level of precision.
\end{desid}

\begin{remark}
	The simplest strategy may be to fix an interval for the value, and use standard criteria of convergence. However, this may be far too restrictive with respect to what we need physically.

	Given that we measure the quantity $f_k(\sigma^k)$ in a finite region $\sigma_k$ with finite precision, there are two things to fix: the confidence interval $\epsilon \in \mathbb{R}$ for the quantity itself and the smallest regions we can measure. As we impose a cutoff in the size of the regions, we do not care whether the approximation holds well at smaller scale (e.g. anything ``higher frequency'' can be ignored, and maybe it doesn't even matter whether higher frequencies diverge).

	We want to define those without reference to a metric/measure on $X$.
\end{remark}

@@

%\section{Insightful failures}

>>>>

\subsection{General definitions}

\begin{defn}
	Let $X$ be a manifold of dimension $n$. A \textbf{k-surface} is a subset $U \subseteq X$ that is a $k$-dimensional manifold (with boundary) with the subspace topology (i.e. regular/embedded submanifold). We note $S^k(X)$, or simply $S^k$, the set of all $k$-surfaces and $S(X) = \bigcup_{i=0}^{n} S^k$ the set of all possible $k$-surfaces of all dimensions.
\end{defn}

\begin{defn}
	A \textbf{k-functional} is a function $f_k : S^k \to \mathbb{R}$ with the following properties:
	\begin{description}
		\item[Linear] $f_k(\sigma_1 \cup \sigma_2) = f_k(\sigma_1) + f_k(\sigma_2)$ for every $\sigma_1, \sigma_2 \in S^k$ such that $\sigma_1 \cap \sigma_2 = \emptyset$
		\item[No contribution from boundary] $f_k(\sigma) = f_k(\sigma \setminus \partial \sigma)$ for every $\sigma \in S^k$
		\item[Commutes with the limit] $\lim\limits_{i \to \infty} f_k(\sigma_i^k) = f_k(\lim\limits_{i \to \infty}\sigma_i^k)$ (limit of a sequence of submain)
	\end{description}
\end{defn}

\subsubsection{Continuity for functionals}

The requirements of commutation with the limit is similar to a definition of continuity in terms of limits. Can we turn that into a topological notion of continuity? Suppose we had a topology on $S^k$, could we simply require the functional to be continuous? How that would constrain $f_k$?

This forms a nice parallel between a map $f : X \to Y$ and the functional $f_k : S^k \to \mathbb{R}$. For both, the open sets are associated to the measurement precision. For $f_k$, the precision of the quantity is an open interval (finite precision) while the precision of the argument is a set of possible $k$ surfaces (i.e. we have an uncertainty on the shape/area).

For example, let $U, V \in S^k$ be two $k$-surfaces without borders such that $U \subset V$. The set $R(U,V) = \{ W \in S^k \, | \, U \subseteq W \subseteq V\}$ is called the \textbf{surface range} between $U$ and $V$. That should be an open set in the topology of $S^k$.

@@

\subsubsection{title}

@@

\subsubsection{Convergence at scale}

We need to understand what is the proper criteria for convergence that correctly reflect the physical approximations we are looking for. The general idea is that we want an approximation that works from a given level of scale upwards, leaving errors at lower scale. This is not how typically things work with standard types of convergence.

Suppose we have a function $f(x)$. Consider now $\hat{f}(x) = f(x) + \epsilon$, where $\epsilon$ can be understood as a small constant error over the whole space. The difference in integral $\int_U \hat{f}(x) - \int_U f(x)dx = \epsilon \Delta x$ increases with the size of the region. In the case the values $x$ and $f(x)$ represent directly the value measured (e.g. $x$ is temperature and $f(x)$ is the length of a metal bar), $\hat{f}$ represents a good approximation. If what we measure directly, however, is the integral (e.g. $U$ is a region and $\int_U f(x)dx$ is the mass in that region), then $\hat{f}$ would give a poor approximation at large scale. For this second case, we need a different method of convergence.

A first attempt could be the following.

\begin{defn}
	Let $X$ be a differentiable manifold. Let $d : X \times X \to [0, +\infty)$ be a metric that induces the topology of $X$. Let $B^k(X, \delta)$ be the set of all $k$-dimensional balls with radius greater than $\delta$. Let $\omega$ be a $k$-form. A sequence of $k$-forms $\{\omega_i\}$ \textbf{converges at scale} to $\omega$ if given $\epsilon, \delta \in \mathbb{R}^+$ we can find $N$ such that $\| \int_U \omega_n - \int_U \omega \| < \epsilon$ for all $n \geq N$ and $U \in B^k(X, \delta)$.
\end{defn}

\begin{defn}
	Let $X$ be a differentiable manifold. Let $d : X \times X \to [0, +\infty)$ be a metric that induces the topology of $X$. Let $B^k(X, \delta)$ be the set of all $k$-dimensional balls with radius greater than $\delta$. Let $f_k$ be a $k$-functional. A sequence of $k$-forms $\{\omega_i\}$ \textbf{converges at scale} to $f_k$ if given $\epsilon, \delta \in \mathbb{R}^+$ we can find $N$ such that $\| \int_U \omega_n - f_k(U) \| < \epsilon$ for all $n \geq N$ and $U \in B^k(X, \delta)$.
\end{defn}

Then we would like to prove the following:

\begin{desid}
	Let $X$ be a differentiable manifold. For every $k$-functional $f_k$ there exists a sequence of $k$-forms $\{\omega_i$\} that converges at scale to $f_k$.
\end{desid}

\section{Graveyard}

\subsection{Coarsening}

This was an attempt to define how to coarsen integrand regions in a way that doesn't depend on a metric.

@@
